---
title: "Computational Musicology Portfolio"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    storyboard: true
    theme:
      heading_font:
        google: 
          family: Rajdhani
          wght: 700
      base_font:
        google: Fira Sans
      code_font:
        google: Fira Mono
      bg: "#FFFFFF"
      fg: "#A00707" 
      primary: "#A00707"
      secondary: "#39d7b8"
      success: "#39d7b8"
      danger: "#fa5577"
      warning: "#ffb14c"
      info: "#0cc7f1"
---

```{r setup, include=FALSE}
library(flexdashboard)
library(spotifyr)
library(tidyverse)
library(compmus)
library(patchwork)
library(recipes)
library(dplyr)
library(ggraph)
library(ggdendro)
library(heatmaply)
library(tidymodels)
library(purrr)
library(viridis)
library(class)
library(caTools)
library(cluster)


get_conf_mat <- function(fit) {
  outcome <- .get_tune_outcome_names(fit)
  fit |> 
    collect_predictions() |> 
    conf_mat(truth = outcome, estimate = .pred_class)
}  

get_pr <- function(fit) {
  fit |> 
    conf_mat_resampled() |> 
    group_by(Prediction) |> mutate(precision = Freq / sum(Freq)) |> 
    group_by(Truth) |> mutate(recall = Freq / sum(Freq)) |> 
    ungroup() |> filter(Prediction == Truth) |> 
    select(class = Prediction, precision, recall)
}

```
```{r include=FALSE}
povindie <- get_playlist_audio_features("", "6xjecZpyEPH6XdVyBeWUP5")
clasindie <- get_playlist_audio_features("", "37i9dQZF1DXaLaoaQqrfyO")
```

Corpus description 
========================

### Introduction

In the recent Wrapped, Spotify introduced a new genre called pov: indie, which gained much traction throughout various internet communities, as users' tastes were being described with that term. The repeated question among all those groups rang: "What is pov: indie?".

This analysis will focus on comparing pov: indie to what is generally considered classic indie in order to establish core differences between the two. Some driving questions include: 

<ul>
    <li>Is pov: indie more of a "vibe" of indie music, or does it classify as a separate and uniquely different genre? If it is more of a "vibe", what describes it?</li>
    <li>If it is found to be reasonably different, does pov: indie have recognizable precursors? Is there a rough year estimate for when it emerged?</li>
    <li>How big is the overlap between the two genres, if any?</li>
</ul>

The analysis will be conducted by taking a representative sample of around 100 tracks out of each genre and applying some comparative measures. The samples are taken from the "Best of POV: Indie" playlist made by volt.fm and the "Indie Pop 2010s" playlist by Spotify. The best course of action would be to use two playlists created by Spotify; however, even though the company has invented the term pov: indie, no official playlist of this genre exists. This begs the question of the quality of the volt.fm creation, which can reasonably be assumed reliable and representative, as volt.fm is a statistics and music discovery tool. Additionally, both playlists have been reviewed "by eye" to double-check if they can be assumed reliable. 

However, the corpus does pose some limitations. One factor that presumably may be problematic is the quite large variance in release years of songs in the pov: indie playlist. To combat the impact of this, a playlist that covers the entirety of the 2010s was chosen. Yet, this may have a reverse impact on the results - this will be checked and necessary adjustments will be made if needed.

When it comes to artists and tracks that may be atypical, Mitski (specifically "Liquid Smooth"), Hozier, and My Chemical Romance in the pov: indie playlist seem like outliers. Those songs, however, may be interesting to analyze in case pov: indie ends up representing more of a "vibe", rather than separate genre qualities. The classic indie playlist seems more balanced in that regard, though it still contains some artists and tracks that could be considered on the verge of genres (Arcade Fire, Mumford & Sons). As far as typical examples go, pov: indie is a bit difficult to assess since the genre is not clearly defined. In the classic indie playlist, almost all entries are very representative, with examples like Clairo, GROUPLOVE, Wallows, Dayglow, and alt-J.

Preliminary assumption, before conducting this research, is that the difference between the two will be barely noticeable on a statistical and musicological level or that pov: indie will be difficult to classify separately as a genre through a set of distinct features. Yet, the distinction between the two will be more significant in a social context, specifically through the lens of "virality" and popularity on short-content platforms like TikTok or Instagram Reels. 

### Corpus playlists

***

<iframe style="border-radius:12px" src="https://open.spotify.com/embed/playlist/6xjecZpyEPH6XdVyBeWUP5?utm_source=generator&theme=0" width="100%" height="352" frameBorder="0" allowfullscreen="" allow="autoplay; clipboard-write; encrypted-media; fullscreen; picture-in-picture" loading="lazy"></iframe>

***

<iframe style="border-radius:12px" src="https://open.spotify.com/embed/playlist/37i9dQZF1DXaLaoaQqrfyO?utm_source=generator&theme=0" width="100%" height="352" frameBorder="0" allowfullscreen="" allow="autoplay; clipboard-write; encrypted-media; fullscreen; picture-in-picture" loading="lazy"></iframe>



Corpus analysis {.storyboard}
====================================

### Playlist-level feature analysis 

```{r include=FALSE}
combo <-
  bind_rows(
    povindie |> mutate(category = "pov: indie"),
    clasindie |> mutate(category = "indie")
  )
```

```{r}
combo |>
  ggplot(aes(x = loudness, y = acousticness, color=category)) +
  geom_point() +
  geom_smooth() +
  scale_color_discrete("Genre:")

combo %>%
  ggplot(aes(x = category, y = acousticness)) +
  geom_boxplot(fill = "lightblue", color = "blue", alpha = 0.7) +
  theme_light() +
  theme(panel.grid.major = element_line(color = "gray", linetype = "dashed"),
        panel.background = element_rect(fill = "white"))
```

Pov: indie acousticness summary:
```{r}
povindie |>
  summarise(
    mean_acousticness = mean(acousticness),
    sd_acousticness = sd(acousticness),
    median_acousticness = median(acousticness),
    mad_acousticness = mad(acousticness)
  )
```

Classic indie acousticness summary:
```{r}
clasindie |>
  summarise(
    mean_acousticness = mean(acousticness),
    sd_acousticness = sd(acousticness),
    median_acousticness = median(acousticness),
    mad_acousticness = mad(acousticness)
  )
```

***

As an initial point of analysis of track-level features acousticness and loudness were chosen. A scatter plot comparing the two genres was created and both groups were compared against each other. Right away the acousticness in pov: indie varies much more significantly than in classic indie. To further confirm this a box plot was created, which shows the quite significant difference between the span as well as the means of acousticness for both of these genres. A summary of the means and medians also confirms this finding.
This indicates that pov: indie is much more likely to be characterized by acoustic performances compared to more classic indie. However, it also encapsulates more variety, which suggests that the value of acousticness alone may not be enough of an indicator to classify a song into one or the other category.

Additionally, by plotting acousticness against loudness it becomes apparent that pov: indie's loudness spans over a larger interval than classic indie. When plotting other playlist features generally the same assumption holds, indicating pov: indie's larger vastness in variety of tracks that can be considered within that genre. These findings so far confirm the preliminary assumptions, that pov: indie will not exemplify largely unified characteristics, and will span over a variety of values.


### Tempo characteristics

```{r}
plot1 <- ggplot(clasindie, aes(x = clasindie$tempo)) +
  geom_histogram() +
  labs(title="Tempo in Classic Indie", x="Tempo (BPM)", y="Count")

plot2 <- ggplot(povindie, aes(x = povindie$tempo)) +
  geom_histogram() +
  labs(title="Tempo in POV: Indie", x="Tempo (BPM)", y="Count")

combined <- plot1 + plot2
combined
```

***

| Tempo                 | Classic indie | Pov: indie |
|-----------------------|---------------|------------|
| minimum               |       71      |      70    | 
| maximum               |      190      |     208    |
| median                |      117      |     123    |
| mean                  |      121      |     122    |

***

The presented histograms have slightly different distribution. The tempo in classic indie seems to have a more normal distribution, slightly skewed towards the left, meaning towards lower BPM. Pov: indie on the other hand strongly favours lower tempo values. However, the values within pov: indie are additionally more broadly distributed, compared to classic indie, with values ranging from below 75 to over 200 BPM. This indicates a preference towards slower songs, which ties with higher values of acousticness observed before. However, the median and means are almost identical for both, with median being 116 and 119 BPM respectively for classic indie and pov: indie, and with mean being 120 and 122 BPM. 


### Tempograms - pov: indie {data-commentary-width=950}

This tempogram represents the song "Bubble Gum" by Clairo, which is the outlier for pov: indie, with the reported tempo of approximately 208 BPM. Interestingly, though, the tempogram seems to indicate an overall tempo of around 100 BPM with a large jump in the last sew seconds, which makes me question the accuracy of the tempo recognizing algorithm used by Spotify. 

For comparison a tempogram of a track representing the mean tempo of pov: indie was also plotted. The song chosen is "As The World Caves In" by Matt Maltese. Here the overall tempo of 123 BPM is much more clearly visible and supported by the tempogram. The song oscillates roughly around that mark, with variations present through the track. 

However, due to the the misclassifications of the overall tempo of Clairo's "Bubble Gum" the usefulness of tempo analysis in the case of a larger sample like this is on quite shaky grounds and should most likely be taken with a grain of salt. 


***

```{r}
bubble_gum <- get_tidy_audio_analysis("3zksbXteOCeSusJ5Xltr3t")
bubble_gum |>
  tempogram(window_size = 8, hop_size = 1, cyclic = TRUE) |>
  ggplot(aes(x = time, y = bpm, fill = power)) +
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +
  labs(title = "Outlier tempogram (Bubble Gum by Clairo)", x = "Time (s)", y = "Tempo (BPM)") +
  theme_classic()
```

***

```{r}
atwci <- get_tidy_audio_analysis("4JE6agBLHGA5TaF6FlqfBD")
atwci |>
  tempogram(window_size = 8, hop_size = 1, cyclic = TRUE) |>
  ggplot(aes(x = time, y = bpm, fill = power)) +
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +
  labs(title = "Outlier tempogram (As The World Caves In by Matt Maltese)", x = "Time (s)", y = "Tempo (BPM)") +
  theme_classic()
```


### Key analysis

```{r}
keys <- c(povindie$key, clasindie$key)
playlists <- c(rep("povindie", length(povindie$key)), rep("clasindie", length(clasindie$key)))

keys_mapped <- factor(keys, levels = 0:11, labels = c("C", "C#", "D", "D#", "E", "F", "F#", "G", "G#", "A", "A#", "B"))

df <- data.frame(keys = keys_mapped, playlists = factor(playlists))

ggplot(df, aes(x = keys, fill = playlists)) +
  geom_bar(position = "dodge", color = "black") +
  labs(x = "Key", y = "Frequency", title = "Histogram of Keys for pov: indie and classic indie") +
  scale_fill_manual(values = c("povindie" = "skyblue", "clasindie" = "lightgreen")) +
  theme_minimal()
```
```{r}
keys_povindie <- povindie$key
keys_clasindie <- clasindie$key
keys_mapped_povindie <- factor(keys_povindie, levels = 0:11, labels = c("C", "C#", "D", "D#", "E", "F", "F#", "G", "G#", "A", "A#", "B"))
keys_mapped_clasindie <- factor(keys_clasindie, levels = 0:11, labels = c("C", "C#", "D", "D#", "E", "F", "F#", "G", "G#", "A", "A#", "B"))

major_counts_povindie <- sum(keys_mapped_povindie %in% c("C", "C#", "D", "D#", "E", "F", "F#", "G", "G#"))
minor_counts_povindie <- sum(keys_mapped_povindie %in% c("A", "A#", "B"))

major_counts_clasindie <- sum(keys_mapped_clasindie %in% c("C", "C#", "D", "D#", "E", "F", "F#", "G", "G#"))
minor_counts_clasindie <- sum(keys_mapped_clasindie %in% c("A", "A#", "B"))

total_keys_povindie <- length(keys_mapped_povindie)
total_keys_clasindie <- length(keys_mapped_clasindie)

proportion_major_povindie <- major_counts_povindie / total_keys_povindie
proportion_minor_povindie <- minor_counts_povindie / total_keys_povindie

proportion_major_clasindie <- major_counts_clasindie / total_keys_clasindie
proportion_minor_clasindie <- minor_counts_clasindie / total_keys_clasindie


cat("Proportion of major keys in pov: indie:", proportion_major_povindie, "\n")
cat("Proportion of major keys in classic indie:", proportion_major_clasindie, "\n")

```


***
The overall distribution of keys across both genres is vastly different. Classic indie exhibits a much larger preference for C and C# as well as B, while pov: indie favours G, A and E. Additionally, the disparity between the use of F# is rather interesting. Both genres, however, exhibit a preference for major keys, with over 75% of all keys being major for both of the genres.


### Chromagrams - classic indie {data-commentary-width=1000}

Chromagrams of classic indie tracks (and outlier and a representative track) used as a point of comparison for pov: indie tracks. 

***

Chroma and timbre features of a representative track form classic indie:

```{r}
clasindie_1 <- get_tidy_audio_analysis("5qJqzpNdEbt92roedFwoDR") |>
  compmus_align(bars, segments) |>
  select(bars) |>
  unnest(bars) |>
  mutate(
    pitches = map(segments, compmus_summarise, pitches, method = "acentre", norm = "manhattan"),
    timbre = map(segments, compmus_summarise, timbre, method = "mean")
  )

plot_clasindie_1 <- bind_rows(
  clasindie_1 |>
    compmus_self_similarity(pitches, "aitchison") |>
    mutate(d = d / max(d), type = "Chroma"),
  clasindie_1 |>
    compmus_self_similarity(timbre, "euclidean") |>
    mutate(d = d / max(d), type = "Timbre")
) |>
  ggplot(
    aes(
      x = xstart + xduration / 2,
      width = xduration,
      y = ystart + yduration / 2,
      height = yduration,
      fill = d
    )
  ) +
  geom_tile() +
  coord_fixed() +
  facet_wrap(~type) +
  scale_fill_viridis_c(guide = "none") +
  theme_classic() +
  labs(x = "", y = "")

plot_clasindie_1
```

*** 

Chroma and timbre features of an outlier of classic indie:

```{r}
clasindie_2 <- get_tidy_audio_analysis("2IqjKEBiz0CdLKdkXhxw84") |>
  compmus_align(bars, segments) |>
  select(bars) |>
  unnest(bars) |>
  mutate(
    pitches = map(segments, compmus_summarise, pitches, method = "acentre", norm = "manhattan"),
    timbre = map(segments, compmus_summarise, timbre, method = "mean")
  )

plot_clasindie_2 <- bind_rows(
  clasindie_2 |>
    compmus_self_similarity(pitches, "aitchison") |>
    mutate(d = d / max(d), type = "Chroma"),
  clasindie_2 |>
    compmus_self_similarity(timbre, "euclidean") |>
    mutate(d = d / max(d), type = "Timbre")
) |>
  ggplot(
    aes(
      x = xstart + xduration / 2,
      width = xduration,
      y = ystart + yduration / 2,
      height = yduration,
      fill = d
    )
  ) +
  geom_tile() +
  coord_fixed() +
  facet_wrap(~type) +
  scale_fill_viridis_c(guide = "none") +
  theme_classic() +
  labs(x = "", y = "")

plot_clasindie_2
```

### Chromagrams - pov: indie {data-commentary-width=950}

These tracks were chosen based on assumptions about what pov: indie is constituted of. The representative track has a repetitive structure, with a clear beat structure. There are, however, two clear changes visible on the timbre plot, where presumably the climax of the song happens. 

For the presumed outlier has a visibly more varied structure, with more changes in structure. Rare clearer patterns are discernible, but overall, the track is much less structured than its representative counterpart. 

This seems to be an indication of a preference towards a more repetitive track structure, with a clearer beat and more constant repetition. This can be confirmed by comparing the chroma and timbre plots of classic indie. For the outlier of classic indie the structure of the track is much more ordered and clear patterns are visible. This is not the case for the representative track - it is much less ordered, as in, clear, repeated patterns are not as apparent.

***

Chroma and timbre features of a representative track from pov: indie:

```{r}
povindie_1 <- get_tidy_audio_analysis("7s5VQqrjBtrBgZL4pEa46S") |>
  compmus_align(bars, segments) |>
  select(bars) |>
  unnest(bars) |>
  mutate(
    pitches = map(segments, compmus_summarise, pitches, method = "acentre", norm = "manhattan"),
    timbre = map(segments, compmus_summarise, timbre, method = "mean")
  )

plot_povindie_1 <- bind_rows(
  povindie_1 |>
    compmus_self_similarity(pitches, "aitchison") |>
    mutate(d = d / max(d), type = "Chroma"),
  povindie_1 |>
    compmus_self_similarity(timbre, "euclidean") |>
    mutate(d = d / max(d), type = "Timbre")
) |>
  ggplot(
    aes(
      x = xstart + xduration / 2,
      width = xduration,
      y = ystart + yduration / 2,
      height = yduration,
      fill = d
    )
  ) +
  geom_tile() +
  coord_fixed() +
  facet_wrap(~type) +
  scale_fill_viridis_c(guide = "none") +
  theme_classic() +
  labs(x = "", y = "")

plot_povindie_1
```

*** 

Chroma and timbre features of an outlier of pov: indie:

```{r}
povindie_2 <- get_tidy_audio_analysis("19W5OTEcQI3ZoRW1HERMyy") |>
  compmus_align(bars, segments) |>
  select(bars) |>
  unnest(bars) |>
  mutate(
    pitches = map(segments, compmus_summarise, pitches, method = "acentre", norm = "manhattan"),
    timbre = map(segments, compmus_summarise, timbre, method = "mean")
  )

plot_povindie_2 <- bind_rows(
  povindie_2 |>
    compmus_self_similarity(pitches, "aitchison") |>
    mutate(d = d / max(d), type = "Chroma"),
  povindie_2 |>
    compmus_self_similarity(timbre, "euclidean") |>
    mutate(d = d / max(d), type = "Timbre")
) |>
  ggplot(
    aes(
      x = xstart + xduration / 2,
      width = xduration,
      y = ystart + yduration / 2,
      height = yduration,
      fill = d
    )
  ) +
  geom_tile() +
  coord_fixed() +
  facet_wrap(~type) +
  scale_fill_viridis_c(guide = "none") +
  theme_classic() +
  labs(x = "", y = "")

plot_povindie_2
```

### Classification - Dendrograms {data-commentary-width=950}

These are the complete dendrograms representing the clustering of respectively pov: indie and classic indie. As visible on the plots, both genres have a vastly different tree representations. Pov: indie has much smaller clusters, especially closer to the root of the tree. It expands much less evenly than the classic indie dendrogram, largely favoring the right side, indicating larger variation and separation within the data set. 

```{r}
pov <-
  get_playlist_audio_features("", "6xjecZpyEPH6XdVyBeWUP5") |>
  add_audio_analysis() |>
  mutate(
    segments = map2(segments, key, compmus_c_transpose),
    pitches =
      map(segments,
        compmus_summarise, pitches,
        method = "mean", norm = "manhattan"
      ),
    timbre =
      map(
        segments,
        compmus_summarise, timbre,
        method = "mean"
      )
  ) |>
  mutate(pitches = map(pitches, compmus_normalise, "clr")) |>
  mutate_at(vars(pitches, timbre), map, bind_rows) |>
  unnest(cols = c(pitches, timbre))

class <-
  get_playlist_audio_features("", "37i9dQZF1DXaLaoaQqrfyO") |>
  add_audio_analysis() |>
  mutate(
    segments = map2(segments, key, compmus_c_transpose),
    pitches =
      map(segments,
        compmus_summarise, pitches,
        method = "mean", norm = "manhattan"
      ),
    timbre =
      map(
        segments,
        compmus_summarise, timbre,
        method = "mean"
      )
  ) |>
  mutate(pitches = map(pitches, compmus_normalise, "clr")) |>
  mutate_at(vars(pitches, timbre), map, bind_rows) |>
  unnest(cols = c(pitches, timbre))
```

```{r}
pov_unique <- pov %>%
  distinct(track.name, .keep_all = TRUE)

pov_juice <- 
  recipe(
    track.name ~
      danceability +
      energy +
      loudness +
      speechiness +
      acousticness +
      instrumentalness +
      liveness +
      valence +
      tempo +
      duration +
      C + `C#|Db` + D + `D#|Eb` +
      E + `F` + `F#|Gb` + G +
      `G#|Ab` + A + `A#|Bb` + B +
      c01 + c02 + c03 + c04 + c05 + c06 +
      c07 + c08 + c09 + c10 + c11 + c12,
    data = pov_unique
  ) %>%
  step_center(all_predictors()) %>%
  step_scale(all_predictors()) %>%
  prep(data = pov_unique) %>%
  bake(new_data = pov_unique)

pov_juice_df <- as.data.frame(pov_juice)
rownames(pov_juice_df) <- pov_unique$track.name


class_unique <- class %>%
  distinct(track.name, .keep_all = TRUE)

class_juice <- 
  recipe(
    track.name ~
      danceability +
      energy +
      loudness +
      speechiness +
      acousticness +
      instrumentalness +
      liveness +
      valence +
      tempo +
      duration +
      C + `C#|Db` + D + `D#|Eb` +
      E + `F` + `F#|Gb` + G +
      `G#|Ab` + A + `A#|Bb` + B +
      c01 + c02 + c03 + c04 + c05 + c06 +
      c07 + c08 + c09 + c10 + c11 + c12,
    data = class_unique
  ) %>%
  step_center(all_predictors()) %>%
  step_scale(all_predictors()) %>%
  prep(data = class_unique) %>%
  bake(new_data = class_unique)

class_juice_df <- as.data.frame(class_juice)
rownames(class_juice_df) <- class_unique$track.name

```

```{r, include=FALSE}
any(is.na(pov_juice_df))
any(is.na(class_juice_df))
```

***

Pov: indie dendrogram:

```{r}
pov_juice_df <- na.omit(pov_juice_df)
class_juice_df <- na.omit(class_juice_df)

pov_dist <- dist(pov_juice_df, method = "euclidean")
class_dist <- dist(class_juice_df, method = "euclidean")

pov_dist |> 
  hclust(method = "complete") |>
  dendro_data() |>
  ggdendrogram()
```

***

Classic indie dendrogram:

```{r}
class_dist |> 
  hclust(method = "complete") |>
  dendro_data() |>
  ggdendrogram()

```

***


### Classification - k-Means Clustering

```{r}
combined_data <- rbind(povindie, clasindie)
set.seed(123)

split_indices <- sample(1:nrow(combined_data), size = 0.7 * nrow(combined_data))
training_data <- combined_data[split_indices, ]
testing_data <- combined_data[-split_indices, ]

X_train <- training_data[, 6:16]
y_train <- ifelse(training_data$playlist_name == "Best of POV: Indie", "Pov: Indie", "Classic Indie")

k <- 2
kmeans_model <- kmeans(X_train, centers = k)
cluster_assignments <- kmeans_model$cluster
cluster_playlist <- ifelse(cluster_assignments == 1, "Pov: Indie", "Classic Indie")
cluster_labels <- data.frame(cluster = cluster_assignments, playlist = cluster_playlist)

pca_result <- prcomp(X_train, scale. = TRUE)
PC1 <- pca_result$x[, 1]
PC2 <- pca_result$x[, 2]

plot_data <- data.frame(PC1, PC2, cluster = as.factor(cluster_assignments))

ggplot(plot_data, aes(x = PC1, y = PC2, color = cluster)) +
  geom_point(size = 3) +
  scale_color_manual(values = c("red", "blue")) +
  labs(title = "K-means Clustering of Training Data") +
  theme_minimal()
```

***

This k-Means clustering approach was modeled to create two clusters based on Spotify API features to classify individual tracks as either part of pov: indie or classic indie. As visible on the plot the end result is far from perfect. The two distinct colour classes labeled 1 and 2 are respectively analogous with pov: indie and classic indie. Though the plot shows slight variation and grouping, it is not enough to say that these two are separate clusters. A large portion of tracks are even plotted basically one atop of another. This showcases that most likely than not these two are not different enough to be considered two separate genres. 


Conclusion
==========================

Let us return to the initial questions and preliminary assumptions made at the start of this research. The goal of this analysis was to answer if pov: indie can be discerned as a separate genre from the broader understanding of classic indie. The driving points were mainly related to assessing the "vibe" of pov: indie and its overlap or the lack of thereof with classic indie. 

The research utilized various techniques, including comparative plot analysis, clustering and other statistical research methods. Through the course of this analysis it became apparent that pov: indie was much more vast and less coherent than classic indie, be it through track-based features or outlier and archetype chroma and timbre analysis. On the other hand at the very core of both these data sets, they were also surprisingly similar, like through the mean tempo, major key preference and other Spotify feature similarities. 

A final point of analysis was performing a classification task using the combined pov: indie and classic indie data set. This was a final indicator of the distinction between the two being too small to guarantee a clear enough separation between the genres. 

This result was rather expected based on a multitude of factors. Firstly, just by judging the two playlists that made up the corpus "by eye", it was apparent that "Indie Pop 2010s" was a much more coherent and had less clear outliers or tracks that could breach genre borders. Whilst "Best of POV: Indie" was rather a headscratcher at times, with songs that, personally, I would not have classified as even close to the same genre, given such task (for example My Chemical Romance, which is renounced as one of the original and most influential emo bands). Secondly, many artists that did in fact seem reasonably placed on the pov: indie playlist were also present on the classic indie playlist (those include: Alex G, girl in red, Clairo, Dayglow etc.). For some of these, even certain tracks were the same on both playlists. On top of all that, pov: indie was also non existent as a genre playlsit made by Spotify. With the vast amount of micro genre playlists and hyper specific mixes that Spotify releases and updates daily, this was a huge red flag that was raised right at the preparation stages of this research. 

Yet still, the main question stands: what is pov: indie and why does it exist? 

Musicalyst describes the genre as "POV: Indie is a music genre distinguished by its independent, do-it-yourself (DIY) approach to production and promotion. The genre is recognizable by its lo-fi sound, unconventional instrumentation, and DIY aesthetic. It has close associations with the punk and indie rock scenes and is considered a more accessible alternative to mainstream music. POV: Indie provides independent artists with an avenue to express themselves creatively without the constraints of the mainstream music industry." However, this sounds awfully similar to what just indie is and historically has been.

I believe that for pov: indie this boils down to the added concept of "virality" and the phenomenon of short-content media, like TikTok. A vast majority of the songs present on the pov: indie playlist were either at one point or another popular on platforms like TikTok or were made by artists whose other songs have gained traction is a similar manner. This leads me to draw a conclusion that pov: indie is rather a specific "vibe" encapsulated, rather than a set of musical characteristics, vastly separate from its origin. Pov: indie seems to have a catchy fragment or repeated beat, which the chroma and timbre analysis confirms, that can easily become popular by grabbing one's attention or making for a good lip syncing song. An additionally interesting and important point of analysis should be the lyrics of the songs considered within pov: indie, as it seems those have as much impact as the beat in classification of the tracks.

Overall, even though pov: indie does not seem to be vastly musically different from what is generally considered classic indie, it does seem to have significant enough social ties to media usage and practices to possibly be considered a subgenre or an "over-genre", encapsulating many different genres. Because of that aspect, as much as there is little evidence of a new emergent genre in a strictly musicological sense, pov: indie still seems to be an interesting and important area of research that can be branched out further through more sociological or anthropological research.

